# pre-class


Make sure you commit this often with meaningful messages. 




### Question 1:

Using a loop, print the integers from 1 to 50. 

```{r}

#Initial integer in desired set, add one each iteration
for (i in 1:50) {
  print(i)
  i <- i+1
}

```

### Question 2:

A.  Using a loop, add all the integers between 0 and 1000.


```{r}
#Set initial value for summing value (x)
#Set interval for counting value (i)
#Add counting value to summing value
#Use print(x) and print(i) in loop to check work

x = 0
for (i in 0:1000) {
  x <- x + i
}
print(x)

```

B. Now, add all the EVEN integers between 0 and 1000 (hint: use seq())

```{r}
#Set initial values
#Set interval using seq
#Add counting value to summing value
#Use print(x) and print(i) in loop to check work

y = 0
i = 0
for (i in seq(0,1000,2)) {
  y <- y + i
}
print(y)

```

C. Now, repeat A and B WITHOUT using a loop.

```{r}
#Use range to sum
A. 
sum(0:1000)

#Use sequence to sum
B. 
sum(seq(0,1000,2))
```

### Question 3:

Here is a dataframe of survey data containing 5 questions :

```{r, eval=FALSE}
survey <- data.frame(
                     "participant" = c(1, 2, 3, 4, 5, 6),
                     "q1" = c(5, 3, 2, 7, 11, 0),
                     "q2" = c(4, 2, 2, 5, -10, 99),
                     "q3" = c(-4, -3, 4, 2, 9, 10),
                     "q4" = c(-30, 5, 2, 23, 4, 2),
                     "q5" = c(88, 4, -20, 2, 4, 2)
                     )
```
The response to each question should be an integer between 1 and 5. Obviously, we have some bad values in the dataframe. The goal of this problem is to fix them.

A. Using a loop, create a new dataframe called survey.clean where all the invalid values (those that are not integers between 1 and 5) are set to NA.

```{r}

#Create copy of survey in survey.clean
#Iterate through range of rows while...
#Iterating through range of columns
#Conditional statement: if specified element is between 1 and 5 inclusive
#Then keep specified element
#Otherwise, replace specified element with NA

survey.clean <- survey
for (i in 1:nrow(survey.clean)) {
  for (j in 2:ncol(survey.clean)) {
    if (survey[i,j] > 1 & survey[i,j] < 5) {
         survey.clean[i,j] = survey.clean[i,j]
    } else {
         survey.clean[i,j] = NA
      }
  }
}
survey.clean

```



B. Now, again using a loop, add a new column to the dataframe called “invalid.answers” that indicates, for each participant, how many bad answers they gave.

```{r}

#Load dplyr package

#Create column of 0's named invalid.answers in survey.clean
#Loop through each element, keeping rows grouped
#Each element in new column equals the sum of the elements in the row that are NA
#Remember is.na returns a 1 if TRUE


library(dplyr)

survey.clean$invalid.answers <- 0
      for (i in 1:nrow(survey.clean)) {
        survey.clean$invalid.answers[i] = sum(is.na(survey.clean[i,])) 
  }
         
survey.clean


```