---
title: "pre-class-03"
output: html_document
---
# pre-class


Make sure you commit this often with meaningful messages. 


### Getting Started


We will work with the dataset called [gapminder](https://github.com/jennybc/gapminder), this is a cleaned up version from [Gapminder Data](http://www.gapminder.org/data/). Gapminder contains a lot of great data on all of the nations of the world. We first need to install the gapminder package in R. 

```{r}
install.packages("gapminder")
```



```{r}
library(dplyr)
library(gapminder)
gapminder
```



### Pre-Class Problems

Use **dplyr** functions to address the following questions:

1. How many unique countries are represented per continent?

```{r}
#regroup to prep for summarize. summarize, count by country
gapminder %>%
  group_by(continent) %>%
  summarize(n_distinct(country))

```
The number of unique countries by continent are as follows: Africa: 52, Americas: 25, Asia: 33, Europe: 30, Oceania: 2. 


2. Which European nation had the lowest GDP per capita in 1997? 

```{r}
#filter by continent and year. arrange in ascending order. Choose first observation. 
gapminder %>%
  filter(continent == "Europe", year == "1997") %>%
  arrange(gdpPercap)
  select
  
```
Albania 


3. According to the data available, what was the average life expectancy across each continent in the 1980s?
```{r}
#filter for desired range of years. Group by continent to set up summarize. Summarize using mean on lifeExp column. 
gapminder %>%
  filter(year >= 1980 & year <= 1989) %>%
  group_by(continent) %>%
  summarize(avg_life_80s = mean(lifeExp))

```
Africa: 52.5;  Americas: 67.2;   Asia:63.7;   Europe: 73.2;   Oceania:74.8


4. What 5 countries have the highest total GDP over all years combined?
```{r}
#Group by country to prime data for summarize. Summarize, add all gdp's per country, create new variable for that data frame. Arrange in descending order. Choose top 5 observations. 
gapminder %>%
  group_by(country) %>% 
  summarize(combined_gdp = sum(gdpPercap)) %>%
  arrange(desc(combined_gdp))
    
```
Kuwait, Switzerland, Norway, United States, Canada

5. What countries and years had life expectancies of _at least_ 80 years? _N.b. only output the columns of interest: country, life expectancy and year (in that order)._
```{r}
gapminder %>%
  filter(lifeExp >= 80) %>%
  select(country, lifeExp, year)
    
```
 1 Australia           80.4  2002
 2 Australia           81.2  2007
 3 Canada              80.7  2007
 4 France              80.7  2007
 5 Hong Kong, China    80    1997
 6 Hong Kong, China    81.5  2002
 7 Hong Kong, China    82.2  2007
 8 Iceland             80.5  2002
 9 Iceland             81.8  2007
10 Israel              80.7  2007
 ... with 12 more rows


6. What 10 countries have the strongest correlation (in either direction) between life expectancy and per capita GDP?
```{r}

    
```

7. Which combinations of continent (besides Asia) and year have the highest average population across all countries? _N.b. your output should include all results sorted by highest average population_. With what you already know, this one may stump you. See [this Q&A](http://stackoverflow.com/q/27207963/654296) for how to `ungroup` before `arrange`ing. This also [behaves differently in more recent versions of dplyr](https://github.com/hadley/dplyr/releases/tag/v0.5.0).
8. Which three countries have had the most consistent population estimates (i.e. lowest standard deviation) across the years of available data? 
9. Subset **gm** to only include observations from 1992 and store the results as **gm1992**. What kind of object is this?
10. Which observations indicate that the population of a country has *decreased* from the previous year **and** the life expectancy has *increased* from the previous year? See [the vignette on window functions](https://cran.r-project.org/web/packages/dplyr/vignettes/window-functions.html).