---
title: "pre-class-03"
output: html_document
---
# pre-class


Make sure you commit this often with meaningful messages. 


### Getting Started


We will work with the dataset called [gapminder](https://github.com/jennybc/gapminder), this is a cleaned up version from [Gapminder Data](http://www.gapminder.org/data/). Gapminder contains a lot of great data on all of the nations of the world. We first need to install the gapminder package in R. 

```{r}
install.packages("gapminder")
```



```{r}
library(dplyr)
library(gapminder)
gapminder
```



### Pre-Class Problems

Use **dplyr** functions to address the following questions:

1. How many unique countries are represented per continent?

```{r}
#Regroup to prep for summarize. 
#Summarize, count unique observations by country. 
gapminder %>%
  group_by(continent) %>%
  summarize(n_distinct(country))

```
The number of unique countries by continent are as follows: Africa: 52, Americas: 25, Asia: 33, Europe: 30, Oceania: 2. 


2. Which European nation had the lowest GDP per capita in 1997? 

```{r}
#Filter by continent and year. 
#Arrange in ascending order. Choose first observation. 
#Keep only first observation. 
gapminder %>%
  filter(continent == "Europe", year == "1997") %>%
  arrange(gdpPercap) %>%
  head(1)

```
Albania 


3. According to the data available, what was the average life expectancy across each continent in the 1980s?
```{r}
#Filter for desired range of years. 
#Group by continent to set up summarize. 
#Summarize using mean on lifeExp column. 
gapminder %>%
  filter(year >= 1980 & year <= 1989) %>%
  group_by(continent) %>%
  summarize(avg_life_80s = mean(lifeExp))

```
Africa: 52.5;  Americas: 67.2;   Asia:63.7;   Europe: 73.2;   Oceania:74.8


4. What 5 countries have the highest total GDP over all years combined?
```{r}
#Group by country to prime data for summarize. 
#Summarize, add all gdp's per country, create new variable for that data frame. 
#Arrange in descending order. Choose top 5 observations. 
#Keep only first 5 observations. 
gapminder %>%
  group_by(country) %>% 
  summarize(combined_gdp = sum(gdpPercap)) %>%
  arrange(desc(combined_gdp)) %>%
  head(5)
    
```
Kuwait, Switzerland, Norway, United States, Canada

5. What countries and years had life expectancies of _at least_ 80 years? _N.b. only output the columns of interest: country, life expectancy and year (in that order)._
```{r}
#Filter for desired life expectancies. 
#Keep only columns of interest. 
gapminder %>%
  filter(lifeExp >= 80) %>%
  select(country, lifeExp, year)
    
```
Australia           80.4  2002; 
Australia           81.2  2007; 
Canada              80.7  2007; 
France              80.7  2007; 
Hong Kong, China    80    1997; 
Hong Kong, China    81.5  2002; 
Hong Kong, China    82.2  2007; 
Iceland             80.5  2002; 
Iceland             81.8  2007; 
Israel              80.7  2007; 
 ... with 12 more rows


6. What 10 countries have the strongest correlation (in either direction) between life expectancy and per capita GDP?
```{r}
#Group by variable of interest to prime data for summary. 
#Summarize and create correlation variable by using the correlation function. 
#Arrange in descending order for highest correlation. 
#Keep only top 10 hits. 
gapminder %>%
  group_by(country) %>%
  summarize(life_gdp_corr = cor(gdpPercap, lifeExp)) %>%
  arrange(desc(life_gdp_corr)) %>%
  head(10)
    
```
France                 0.996; 
Austria                0.993; 
Belgium                0.993; 
Norway                 0.992; 
Oman                   0.991; 
United Kingdom         0.990; 
Italy                  0.990; 
Israel                 0.988; 
Denmark                0.987; 
Australia              0.986; 


7. Which combinations of continent (besides Asia) and year have the highest average population across all countries? _N.b. your output should include all results sorted by highest average population_. With what you already know, this one may stump you. See [this Q&A](http://stackoverflow.com/q/27207963/654296) for how to `ungroup` before `arrange`ing. This also [behaves differently in more recent versions of dplyr](https://github.com/hadley/dplyr/releases/tag/v0.5.0).
```{r}
#Only keep countries not in Asia. 
#Group by to prime data for summary. 
#Summarize to calculate average of population for all countries in each continent. 
#Arrange those averages. 
gapminder %>%
  filter(country != "Asia") %>%
  group_by(continent, year) %>%
  summarize(avg_pop = mean(pop)) %>%
# ungroup(avg_pop)%>% if necessary 
  arrange(avg_pop)
    
```
Africa     1952 4570010; 
Africa     1957 5093033; 
Oceania    1952 5343003; 
Africa     1962 5702247; 
Oceania    1957 5970988; 
Africa     1967 6447875; 
Oceania    1962 6641759; 
Oceania    1967 7300207; 
Africa     1972 7305376; 
Oceania    1972 8053050; 
... with 50 more rows

8. Which three countries have had the most consistent population estimates (i.e. lowest standard deviation) across the years of available data? 
```{r}
#Group by column of interest for summary. 
#Summarize, define new variable, and use standard deviation function on population. 
#Arrange in ascending order. 
#Return first 3 observations. 
gapminder %>%
  group_by(country) %>%
  summarize(st_dev = sd(pop)) %>%
  arrange(st_dev)%>%
  head(3)
```
Sao Tome and Principe 45906; 
Iceland               48542; 
Montenegro            99738; 


9. Subset **gm** to only include observations from 1992 and store the results as **gm1992**. What kind of object is this?
```{r}
#Filter for year of interest and set to new variable. 
#Function to determine class of object. 
gapminder %>%
  filter(year == "1992") -> gm1992

class(gm1992)
```
gm1992 is a tibble dataframe


10. Which observations indicate that the population of a country has *decreased* from the previous year **and** the life expectancy has *increased* from the previous year? See [the vignette on window functions](https://cran.r-project.org/web/packages/dplyr/vignettes/window-functions.html).
```{r}
#Keep observations based if population has decreased from previous year, lag function compares value from previous year. 
#Add new column life_expect that determines if life expectancy has increased from previous year. 
#Only keep observations where this is true. 
#Arrange population in ascending order. 
gapminder %>%
  filter(pop < lag(pop)) %>%
  mutate(life_expect = lifeExp > lag(lifeExp)) %>%
  filter(life_expect == TRUE)%>%
  arrange(pop)

```


