# pre-class


Make sure you commit this often with meaningful messages. 



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


1. Read the source code for each of the following three functions, puzzle out what they do, and then brainstorm better names.

```{r}

string_a <- c(1,2,3,4,5)
prefix_a <- 3

string_b <- c("This", "is", "an", "amazing", "string")
prefix_b <- "is"

#Iterates through each value in string, if equal to prefix, print TRUE, if not, print FALSE
string_contains_prefix <- function(string, prefix) {
  substr(string, 1, nchar(prefix)) == prefix
}

#Test:
string_contains_prefix(string_b, prefix_b)

```

```{r}

a <- c(1,2,3,5,43,56,7,7,4)
  
  
#Removes last element in sequence if sequence contains more than one element
remove_last_element <- function(x) {
  if (length(x) <= 1) return(NULL)
  x[-length(x)]
}

#Test:
remove_last_element(a)

```

```{r}

b <- c(1,2,3,4,5,6,7,8,9)
c <- c(4,5,6,6,7,8)

#Repeats second list until it is the length of the first list
repeat_equal_lengths <- function(x, y) {
  rep(y, length.out = length(x))
}

#Test
repeat_equal_lengths(b,c)

```

2. Compare and contrast rnorm() and MASS::mvrnorm(). How could you make them more consistent?
```{r}


```
3. Use `lapply()` and an anonymous function to find the coefficient of variation (the standard deviation divided by the mean) for all columns in the mtcars dataset. 
```{r}


```

4. Use vapply() to:
    a. Compute the standard deviation of every column in a numeric data frame.
```{r}


```
    b. Compute the standard deviation of every numeric column in a mixed data frame. (Hint: you'll need to use vapply() twice.)
```{r}


```