# pre-class


Make sure you commit this often with meaningful messages. 



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```








Standardizing a variable means subtracting the mean, and then dividing by the standard deviation. Let's use a loop to standardize the numeric columns in the [Western Collaborative Group Study](https://clinicaltrials.gov/ct2/show/NCT00005174). This study began in 1960 with 3154 men ages 39-59, who were employed in one of 11 California based companies. They were followed until 1969 during this time, 257 of these men developed coronary heart disease (CHD). You can read this data in with the code below. You can access this dataset with the following code:

```{R}

suppressMessages(library(foreign))
wcgs <- read.dta("https://drive.google.com/uc?export=download&id=0B8CsRLdwqzbzYWxfN3ExQllBQkU")
```

The data has the following variables:



WCGS has the following variables:

-----------------------------------------------------------
Name    Description
------- -------------------------------------------
id      Subject identification number

age     Age in years

height  Height in inches

weight  Weight in lbs.

sbp     Systolic blood pressure in mm 

dbp     Diastolic blood pressure in mm Hg

chol    Fasting serum cholesterol in mm 

behpat  Behavior

  1       A1

  2       A2

  3       B3

  4       B4

ncigs   Cigarettes per day

dibpat  Behavior

1       type A

2       type B

chd69   Coronary heart disease

1       Yes

0       no

typechd Type of CHD

1       myocardial infarction or death

2       silent myocardial infarction

3       angina perctoris

time169 Time of CHD event or end of follow-up

arcus   Arcus senilis

0       absent

1       present

bmi     Body Mass Index
-----------------------------------------------------------




### Question 1: Standardize Function

A. Create a function called standardize.me() that takes a numeric vector as an argument, and returns the standardized version of the vector. 
```{r}
#Define function and argument
  #To standardize variable: subtract mean from variable then take standard deviation

standardize.me <- function(num_vec) {
  (num_vec-mean(num_vec))/sd(num_vec)

}

#test below
#standardize.me(c(10,20,30))

```
B. Assign all the numeric columns of the original WCGS dataset to a new dataset called WCGS.new.
```{r}
#Define new variable, use dplyr to select_if(data, condition)

WCGS.new <- dplyr::select_if(wcgs, is.numeric)

```
C. Using a loop and your new function, standardize all the variables WCGS.new dataset.
```{r}
#For element in all the elements in data
  #Take the standardization of each element, define it back into each element in dataframe

for (i in 1:length(WCGS.new)) {
  WCGS.new[i] <- standardize.me(WCGS.new[i])
}

#result below
#WCGS.new

```
D. What should the mean and standard deviation of all your new standardized variables be? Test your prediction by running a loop
```{r}
#The mean should be 0 and the sd should be 1. Use example c(1,2,3) with standardization = -1,0,1. mean=0, sd=1. 

#For each column in data
#Calculate the mean of all the elements in each column, removing NA's
#Calculate the sd of all the elements in each column, removing NA's

for (i in 1:ncol(WCGS.new)) {
  WCGS.new.mean <- mean(WCGS.new[,i], na.rm=TRUE)
  WCGS.new.sd <- sd(WCGS.new[,i], na.rm=TRUE)
}

#Test below
#WCGS.new.mean
#WCGS.new.sd


```




### Question 2: Looping to Calculate

A. Using a loop, calculate the mean weight of the subjects separated by the type of CHD they have.
```{r}
#Select columns of interest for readability
#Initialize variables

#For weight elements in column wt_typ
  #if the corresponding typchd69 value is no CHD 
  #then sum these values together, save as variable

  #if the corresponding typchd69 value is MI or SD 
  #then sum these values together, save as variable

  #if the corresponding typchd69 value is silent MI 
  #then sum these values together, save as variable

  #if the corresponding typchd69 value is any other value 
  #then sum these values together, save as variable

wt_typ <- dplyr::select(wcgs,"weight", "typchd69")
no_CHD <- 0
MI_or_SD <- 0
silent_MI <- 0
angina <- 0

for (i in wt_typ) {
  if("typchd69" == "no CHD") {
    no_CHD <- sum(i)
  }
  else if("typchd69" == "MI or SD"){
    MI_or_SD <- sum(i)
  } 
  else if("typchd69" == "silent MI"){
    silent_MI <- sum(i)
  }
  else {
    angina <- sum(i)
  }

}

#Test below
#no_CHD
#MI_or_SD
#silent_MI
#angina

```
B. Now do the same thing, but now don't use a loop
```{r}

#Load dplyr library
#Begin piping with dataset
  #Group by the type of coronary heart disease to set up summary
  #Summarize data, extract the mean of the weights for each type of CHD

library(dplyr)

wcgs %>%
  group_by(typchd69) %>%
  summarise(mean(weight))

```

```{r}

#Misc notes
#dplyr::group_by(wcgs,"typechd69")
#wt_typ <- dplyr::select(wcgs,"weight", "typchd69")
#sorted <- dplyr::group_by(wt_typ,"typchd69")
#wcgs[,"weight"]

```
